---
title: "portfolio-2-tidy-tues-bees"
author: "Elayna Seago"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

My goals for my bee colony portfolio project are to:
1) get practice using Tidy Tuesday data
2) determine what 10 states/times experienced the greatest loss in colony numbers over a 3 month period. 
3) make a graph for the stressors in each of those states in their worst year.

## the cleaning of this data was done using the code provided by tidy tuesday bee data
```{r bee-data}

colony <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/colony.csv')
stressor <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/stressor.csv')


```


```{r libraries}
library(rvest)
library(tidyverse)
library(fs)
```



```{r}
# download site
#url <- "https://usda.library.cornell.edu/concern/publications/rn301137d?locale=en"


```




### I wanted to see where the greatest 1 year losses in bee colonies were. I found that New Mexico had the worst 3 month period and also the 10th and 11th worst 3 month periods.

```{r get-top-ten}
max_colony_loss <- colony %>%              
  arrange(desc(colony_lost_pct)) %>% 
  slice(1:10)

max_colony_loss
```

#fix this graph-only 1 new mexico
```{r make-graph}
max_colony_loss%>% 
  ggplot(aes(x = reorder(state , -colony_lost_pct) , y = colony_lost_pct)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) 
```

After making this graph I realized I need to glue state, month, and year together because otherwise New Mexico is very misleading.

```{r glue}
library(glue)

max_colony_loss %>% 
  mutate(
    time_place = glue("{year} {months} {state}")
  ) %>% 
  ggplot(aes(x = reorder(time_place , -colony_lost_pct) , y = colony_lost_pct)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(x = "Location and Three Month Period" , y = "Percent of Hives Lost" , title = "Percent Bee Hives Lost" , subtitle = "In the 10 Locations and 3 Month Periods with the Greatest Loss")
```

### Goal number 3
I think it will be helpful to join the two data sets for this portion.
```{r join-data-sets}
colonies_stressors <- full_join(colony, stressor, by =  c("year" , "months" , "state"))

colonies_stressors

```
#re-do the filter thing I did before to get the top 10 worst sections and do the glueing-sliced for 1:60 because it is the long version of data
# initially I thought I had made some sort of mistake because the pct stressors adds to above 100% in some cases, but then I realized one hive can have multiple stressors. So I think this is actually correct after all. however, this may make the graph confusing to others. It would be nice if there was a different stressor option that said "multiple stressors" or something.
```{r stressor-graph}
max_colony_loss_stress <- colonies_stressors %>%              
  arrange(desc(colony_lost_pct)) %>% 
  slice(1:60)

max_colony_loss_stress %>% 
  mutate(
    time_place = glue("{year} {months} {state}")
  ) %>% 
  ggplot(aes(x = time_place , y = stress_pct , fill = stressor)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(x = "Location and Three Month Period" , y = "Percent of Hives With Each Stressor" , title = "Bee Hive Stressors in the 10 States with the Greatest Percent Hive Loss" , subtitle = "In the 3 Month Period with the Greatest Loss")
```


```{r understanding_stress_pct_variable}

max_colony_loss_stress

```



